Reduced GW invariants from cuspidal curves

### 2017-12-11 draft.tex
Introduction and section 1

\newcommand{\M}[4]{\overline{\mathcal M}_{#1,#2}(#3,#4)}
\newcommand{\Mone}[3]{\overline{\mathcal M}^{(1)}_{#1}(#2,#3)}
\newcommand{\PP}{\mathbb P}
\newcommand{\OO}{\mathcal O}
\renewcommand{\to}{\rightarrow}
\newcommand{\pr}{\rm{pr}}
\newcommand{\Aaff}{\mathbb A}
\newcommand{\A}{\mathcal A}
\newcommand{\B}{\mathcal B}
\newcommand{\X}{\pazocal X}
\newcommand{\XP}{\pazocal{X\!P}}
\newcommand{\tXP}{\widetilde{\pazocal{X\!P}}}
\renewcommand{\L}{\mathcal L}
\newcommand{\hL}{\hat{\mathcal L}}
\newcommand{\C}{\mathfrak C}
\newcommand{\cC}{\mathcal C}
\newcommand{\Z}{\pazocal Z}
\newcommand{\Zp}{\pazocal Z^p}
\newcommand{\tZ}{\widetilde{\pazocal Z}}
\newcommand{\tZp}{\widetilde{\pazocal Z}^p}
\newcommand{\MM}{\mathfrak M}
\newcommand{\pP}{\mathfrak P}
\newcommand{\hP}{\hat{\mathfrak P}}
\newcommand{\hM}{\hat{\mathfrak M}}
\newcommand{\oM}{\overline{\pazocal M}}
\newcommand{\tM}{\widetilde{\pazocal M}}
\newcommand{\R}{\operatorname{R}}
\newcommand{\Gm}{\mathbb{G}_{\rm{m}}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\PtoM}{\rho}
\newcommand{\vir}[1]{[#1]^{\rm{vir}}}
\newcommand{\virloc}[1]{[#1]^{\rm{vir}}_{\rm{loc}}}

@@

The moduli stack of stable maps to projective space is a fundamental object in Gromov-Witten theory, since any moduli space of maps to a (smooth) projective variety is cut inside that of the ambient projective space by a set of induced equations. Yet its structure in higher genus is still largely mysterious.

Recall that, in genus zero, $\M{0}{n}{\PP^r}{d}$ is a smooth stack of the expected dimension $\operatorname{vdim}=r-3+d(r+1)+n$ (we see that it is unobstructed from the fact that $H^1(\PP^1,f^*T_{\PP^r})=0$ for any map $f\colon\PP^1\to\PP^r$). Furthermore, for any split vector bundle $E=\oplus_{i} \mathcal O_{\PP^r}(l_i)$, the sheaf $\pi_*f^*E$ (where $\pi$ is the structure map of the universal curve and $f$ the universal stable map) is actually a vector bundle on $\M{0}{n}{\PP^r}{d}$.
\bcd
\mathcal C_{0,n}(\PP^r,d)\ar[r,"f"]\ar[d,"\pi"] & \PP^r\times \M{0}{n}{\PP^r}{d} \\
\M{0}{n}{\PP^r}{d} &
\ecd

It is a matter of functoriality of virtual fundamental classes \cite{KON} \cite{CKL} \cite{KKP} that, if $i\colon X\to \PP^r$ is a complete intersection of degree $(l_1,\ldots,l_k)$ (if $X$ is cut by a generic section $s\in H^0(\PP^r,E)$, then the induced section $\tilde s=\pi_*f^*(s)$ is also generic) and we denote by $j$ the corresponding inclusion
$$\bigsqcup_{\beta\in A_1(X)\colon i_*(\beta)=d}\M{0}{n}{X}{\beta}\hookrightarrow\M{0}{n}{\PP^r}{d}$$
then the following equality, expected by the interpretation of Chern classes as zero loci, holds:
$$j_*[\bigsqcup_{\beta\in A_1(X)\colon i_*(\beta)=d}\M{0}{n}{X}{\beta}]^{\operatorname{vir}}=c_{\operatorname{top}}(E)\cap[\M{0}{n}{\PP^r}{d}].$$
This so-called ``hyperplane property'' allows us to compute GW invariants of complete intersections as twisted invariants on projective space.
The situation in higher genera is much more intricated: the moduli space $\M{g}{n}{\PP^r}{d}$ is neither irreducible, nor pure dimensional; the sheaf $\pi_*f^*\mathcal O_{\PP^r}(l)$ is not a vector bundle. One of the reason for this (in fact the only reason in genus one) is that there are maps from reducible curves with a genus $g$ core that is contracted and some rational tails that carry all the degree; if $f\colon C\to \PP^r$ is such a map, then $H^1(C,f^*\mathcal O_{\PP^r}(1))\neq 0$, as can be easily deduced from the normalisation sequence.
\section{Genus one stable maps to $\PP^r$ - Equations, components and alternate compactifications}

\subsection{Local equations and components}

The geometry of the genus one moduli space of Kontsevich' stable maps to projective space $\M{1}{n}{\PP^r}{d}$ has been widely studied since \cite{Vre} and is by now well understood. Assume $d>0$. Let $E$ be a smooth genus one curve with a non-constant morphism $f\colon E\to \PP^r$; then $H^1(E,f^*T_{\PP^r})=0$ by Riemann-Roch, hence we may define the \emph{main component} of $\M{1}{n}{\PP^r}{d}$ as the closure of the locus of maps the source of which is a smooth genus one curve. It is irreducible of the expected dimension $\operatorname{vdim}=d(r+1)+n$. On the other hand, for every positive integer $k$ and partitions $\lambda$ of $d$ into $k$ (positive) parts and $\mu$ of $n$ into $k$ (positive) parts, consider the boundary component $D_{\lambda,\mu}(\PP^r,d)$ that is the closure of the locus where (i) the source curve is obtained by gluing a smooth elliptic curve with $k$ many $\PP^1$'s (as rational tails), (ii) the map contracts the core elliptic curve to a point, and (iii) the rational tail $R_i$ has degree $\lambda_i$ and $\mu_i$ many marked points. In fact $D_{\lambda,\mu}(\PP^r,d)$ is the image of the gluing morphism from the fiber product
$$\M{1}{k}{\PP^r}{0}\times_{\PP^r}\prod_{i=1}^k\M{0}{\mu_i+1}{\PP^r}{\lambda_i}.$$

\begin{prop}\label{prop:components}
\emph{(1)} The set of irreducible components of $\M{1}{n}{\PP^r}{d}$ is in bijection with $\{(\lambda\vdash d,\mu\vdash n)|\ \#(\text{parts})(\lambda)=\#(\text{parts})(\mu)\}\cup\{\circ=\operatorname{main}\}$, i.e.
$$\M{1}{n}{\PP^r}{d}=\M{1}{n}{\PP^r}{d}^\circ\cup\bigcup_{\lambda,\mu}D_{\lambda,\mu}(\PP^r,d).$$

\emph{(2)} Let $(E {}_{p_1,\ldots,p_k}\sqcup_{q_1,\ldots,q_k}\bigsqcup_{i=1}^k R_i,f)$ be a degenerate stable map with $k$ rational tails; then it lies in the boundary of the main component if and only if $\operatorname{d}\!f(T_{q_i}R_i)$ are \emph{linearly dependent} in $T_{f(E)}\PP^r$.
\end{prop}

In order to understand how the Gromov-Witten invariants differ from the naive expectation of counting smooth embedded genus $g$ curves in the given homology class, one needs to have a good understanding of the virtual fundamental class and of its splitting on the various components. 
This has led to the study of explicit local equation for the moduli space of stable maps in a smooth (over $\mathfrak M_{1,n}$) ambient space \cite{HL}, which we shall now briefly explain. Recall that a map from a scheme $C$ to projective $r$-space is the same as the datum of a line bundle $L$ on $C$ together with $r+1$ global sections in $H^0(C,L)$ that generate the line bundle at every point of $C$. Then it seems natural to embed the space of stable maps as an open inside $\pi_*\mathcal L^{\oplus r+1}$ on the universal Picard stack $\mathfrak{P}_{1,n}$. The problem is that this cone is not smooth, due to the jumps in fiber dimension at the boundary.
%(besides there is a small issue due to the $\mathbb G_m$ automorphism group of objects of $\mathfrak{P}_{1,n}$, which can be avoided by dishomogenising).

Let $[f\colon C\to \PP^r]$ be a point of $\M{1}{n}{\PP^r}{d}$; we may fix homogeneous coordinates on $\PP^r$ in a way that $D_0:=f^*\{X_0=0\}$ is a $d$-uple of smooth points on $C$ (i.e. \'{e}tale on the base) (this property will hold in a neighbourhood of $[f]$). This gives a map from (an \'{e}tale chart of) $\M{1}{n}{\PP^r}{d}$ to (an \'{e}tale chart of) the moduli space of pre-stable curves with a degree $d$ divisor $\mathfrak M_{1,n}^{\operatorname{div}=d}$, that we may assume to be landing in the locus where the divisor is composed by $d$ smooth distinct points (notice this locus is smooth over $\mathfrak M_{1,n}$ by the deformation theory of smooth subschemes). A morphism to $\PP^r$ shall now be thought of as a curve-divisor pair $(C,D)$ together with $r$ global sections of $\mathcal O_C(D)$: the morphism can be written as $[1:u_1:\ldots:u_r]$, where $1$ is the image of the constant global section along the given map $\mathcal O_C\to\mathcal O_C(D)$ (i.e. we have taken dishomogenised local coordinates on $\PP^r$).

Furthermore, \'{e}tale locally on $\M{1}{n}{\PP^r}{d}$, we may pick extra sections $\mathcal A$ and $\mathcal B$ of $\mathcal C_{1,n}(\PP^r,d)$ such that (i) they pass through the core elliptic curve, and (ii) they are disjoint smooth points away from $\mathcal D_0$. Then $\pi_*\mathcal L(\mathcal A)$ \emph{is} a vector bundle on $\mathfrak P_{1,n}$ and $\pi_*\mathcal L$ is carved inside it by requiring that the restriction (residue) map $\pi_*\mathcal L(\mathcal A)\to\pi_*\mathcal L(\mathcal A)_{|\mathcal A}$ is zero.

\begin{prop}
\emph{(1)} \'{E}tale locally, there is a locally closed embedding of $\M{1}{n}{\PP^r}{d}$ inside the vector bundle $V_1:=\Spec(\pi_*\mathcal L(\A)^{\oplus r})$ over $\mathfrak M_{1,n}^{\operatorname{div}=d}$. It is obtained by imposing the nondegeneracy and stability conditions \emph{(open)} on the vanishing locus of the restriction of the universal section of $\pi_*\mathcal L(\A)^{\oplus r}$ to $\A$ \emph{(closed)}.

\emph{(2)} Let $\mathcal V$ be an (affine local) \'{e}tale chart around a smooth point of $\mathfrak M_{1,n}^{\operatorname{div}=d}$. Then $\pi_*\mathcal L(\A)\cong \pi_*\mathcal L(\A-\B)\oplus \pi_*\mathcal \mathcal L(\A)_{|\B}$ (trivially, $\mathcal O_{\mathcal V}^{\oplus d+1}\cong\mathcal O_{\mathcal V}^{\oplus d}\oplus\mathcal O_{\mathcal V}$) and the restriction-to-$\A$ map is zero on the second factor. Call $\varphi\colon \pi_*\mathcal L(\A-\B)\to\pi_*\mathcal L(\A)_{|\A}$ the induced  restriction to $\A$.

\emph{(3)} If we choose a suitable basis of $\pi_*\mathcal L(\A-\B)$, the restriction-to-$\A$ map can be written in a very explicit form: let $D=\sum_{i=1}^d\delta_i$, then $\pi_*\mathcal O_{\mathfrak C}(\delta_i+\A-\B)$ is a sub-line bundle of $\pi_*\mathcal L(\A-\B)$ and $\varphi=\oplus\varphi_i\colon\oplus_{i=1}^d \pi_*\mathcal O_{\mathfrak C}(\delta_i+\A-\B)\to \pi_*\mathcal L(\A)_{|\A}.$ Furthermore $\varphi_i\colon\mathcal O_\mathcal V\to\mathcal O_\mathcal V$ is given (up to invertibles) by multiplication by $\prod_{q\in[\A,\delta_i]}\zeta_q$, where $\zeta_q$ is the smoothening variable on $\mathfrak M_1$ corresponding to the node $q$, and $q\in[a,\delta_i]$ stands for all the nodes separating $\A$ (i.e. the core elliptic curve) from the point $\delta_i$.
\end{prop}

Given these explicit equations and by carefully examining the proof of \cite[Prop. 4.13]{HL}, we can look back at the problem of determining the boundary of the main component of the space of stable maps. The following characterisation of smoothability was already known to (and partly proven in) \cite[Lem. 5.9]{Vre}.

\begin{prop}\label{smoothability_with_eqn}
Let $[f\colon C\to \PP^r]$ be a degenerate stable map from a ($n$-marked) genus one curve of the form $(E {}_{p_1,\ldots,p_m}\sqcup_{q_1,\ldots,q_m}\bigsqcup_{i=1}^m R_i,f)$, where $E$ is the maximal connected contracted arithmetic genus one subcurve, and $R_i$ are the rational tails (chains of $\PP^1$ along which $f$ has positive degree). The following are equivalent:
\begin{enumerate}
\item $[f]$ is smoothable;
\item $\operatorname{d}\!f(T_{q_i}R_i)$ are linearly dependent in $T_{f(E)}\PP^r$;
\end{enumerate}
\end{prop}
\begin{proof}
 Let us start with the easiest degenerate situation: a contracted elliptic curve joined with a rational tail of degree $d$ at a single node $q$. Equations for the moduli space of maps around such a point look like $\zeta_q\sum_{i=1}^d w_i^j=0$, for $j=1,\ldots,r$, the $w_i$'s being coordinates for $\pi_*\mathcal O_{\mathfrak C}(D+\A-\B)$ in the basis given by $\pi_*\mathcal O_{\mathfrak C}(\delta_i+\A-\B),\ i=1,\ldots,d$. Our point corresponds to a smoothable map if and only if the equations admit a solution with $\zeta_q\neq 0$, hence it must be $\sum_{i=1}^d w_i^j=0$ for every $j$. Taking a coordinate $z$ on the rational tail around the node $q$ (i.e. $\{z=0\}$ corresponds to the node), we see that the $i$-th vector in the chosen basis corresponds to a polynomial vanishing at the node and at every other point $\delta_j,\ j\neq i$; hence that can be written as $e_i(z)=z\prod_{j\neq i}\frac{(z-\delta_j)}{-\delta_j}$ (we chose a convenient normalisation). So the map corresponding to the point of coordinates $(w_i^j)_{i=1,\ldots,d;j=1,\ldots,r}$ can be represented as $[1:\sum_{i=1}^d w_i^1e_i(z):\ldots,\sum_{i=1}^d w_i^re_i(z)]$ on the rational tail; differentiating we see that the tangent vector at the node is sent to $(\sum_{i=1}^d w_i^1,\ldots,\sum_{i=1}^d w_i^r)$ (in affine coordinates), hence smoothability is equivalent to the image of the tangent vector being zero.

>>>>

\newcommand{\M}[4]{\overline{\pazocal M}_{#1,#2}(#3,#4)}
\newcommand{\Mone}[3]{\overline{\pazocal M}^{(1)}_{#1}(#2,#3)}
\newcommand{\PP}{\mathbb P}
\newcommand{\OO}{\mathcal O}
\renewcommand{\to}{\rightarrow}
\newcommand{\pr}{\rm{pr}}
\newcommand{\Aaff}{\mathbb A}
\newcommand{\A}{\mathcal A}
\newcommand{\B}{\mathcal B}
\newcommand{\X}{\pazocal X}
\newcommand{\XP}{\pazocal{X\!P}}
\newcommand{\tXP}{\widetilde{\pazocal{X\!P}}}
\renewcommand{\L}{\mathcal L}
\newcommand{\hL}{\hat{\mathcal L}}
\newcommand{\C}{\mathfrak C}
\newcommand{\cC}{\mathcal C}
\newcommand{\Z}{\pazocal Z}
\newcommand{\Zp}{\pazocal Z^p}
\newcommand{\tZ}{\widetilde{\pazocal Z}}
\newcommand{\tZp}{\widetilde{\pazocal Z}^p}
\newcommand{\MM}{\mathfrak M}
\newcommand{\pP}{\mathfrak P}
\newcommand{\hP}{\hat{\mathfrak P}}
\newcommand{\hM}{\hat{\mathfrak M}}
\newcommand{\oM}{\overline{\pazocal M}}
\newcommand{\tM}{\widetilde{\pazocal M}}
\newcommand{\R}{\operatorname{R}}
\newcommand{\Gm}{\mathbb{G}_{\rm{m}}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\PtoM}{\rho}
\newcommand{\vir}[1]{[#1]^{\rm{vir}}}
\newcommand{\virloc}[1]{[#1]^{\rm{vir}}_{\rm{loc}}}

@@

\section*{Introduction}

The moduli stack of stable maps to projective space is a fundamental object in Gromov-Witten theory, since any moduli space of maps to a projective variety is cut inside the former by a number of equations induced by those of $X\subseteq \PP^r$. Yet its structure in higher genus is still largely mysterious.

Recall that in genus zero $\M{0}{n}{\PP^r}{d}$ is a smooth stack of the expected dimension. Furthermore for any split vector bundle $E=\oplus_{i} \mathcal O_{\PP^r}(l_i)$ the sheaf $\pi_*f^*E$ is a vector bundle on $\M{0}{n}{\PP^r}{d}$.
\bcd
\mathcal C_{0,n}(\PP^r,d)\ar[r,"f"]\ar[d,"\pi"] & \PP^r\times \M{0}{n}{\PP^r}{d} \\
\M{0}{n}{\PP^r}{d} &
\ecd

It is a matter of functoriality of virtual fundamental classes \cite{KON} \cite{CKL} \cite{KKP} that, if $i\colon X\hookrightarrow \PP^r$ is a complete intersection of degree $(l_1,\ldots,l_k)$ and we denote by $j$ the corresponding inclusion of moduli spaces of maps,
%$$\bigsqcup_{\beta\in A_1(X)\colon i_*(\beta)=d}\M{0}{n}{X}{\beta}\hookrightarrow\M{0}{n}{\PP^r}{d}$$
then the following equality holds:%, which is expected from the interpretation of Chern classes as zero loci of sections (if $X$ is cut by a generic section $s\in H^0(\PP^r,E)$, then the induced section $\tilde s=\pi_*f^*(s)$ is also generic):
$$j_*[\bigsqcup_{\beta\in A_1(X)\colon i_*(\beta)=d}\M{0}{n}{X}{\beta}]^{\operatorname{vir}}=c_{\operatorname{top}}(E)\cap[\M{0}{n}{\PP^r}{d}].$$
This so-called ``hyperplane property'' allows us to compute GW invariants of complete intersections as twisted invariants of projective space, which enables us to perform computations by localisation.

The situation in higher genera is much more intricate: the moduli space $\M{g}{n}{\PP^r}{d}$ is neither irreducible, nor pure dimensional, and the sheaf $\pi_*f^*\mathcal O_{\PP^r}(l)$ is not a vector bundle. The enumerative meaning of the invariants is affected by the appearance of unexpected components, representing degenerate contributions from curves of lower genera.

The situation in genus $1$ is better understood: the geometry of the moduli space of genus one stable maps to projective space $\M{1}{n}{\PP^r}{d}$ has been widely studied since \cite{Vre}. Contrary to higher genus, the closure of the locus of maps the source of which is a smooth elliptic curve is an irreducible component of $\M{1}{n}{\PP^r}{d}$, which is usually called the \emph{main component}.
%One of the reasons for this (in fact the only reason in genus one) is that there are maps from reducible curves with a genus $g$ core that is contracted and some rational tails that carry all the degree; if $f\colon C\to \PP^r$ is such a map, then $H^1(C,f^*\mathcal O_{\PP^r}(1))\neq 0$, as can be easily seen from the normalisation sequence.

In an attempt to extract the enumerative information hidden within Kontsevich's moduli space, Vakil and Zinger managed to desingularise the main component by performing a sequence of blow-ups on an underlying space of weighted curves \cite{VZpreview} \cite{VZ} (see also \cite{HL}). The end result $\widetilde{\pazocal M}_{1,n}(\PP^r,d)^{\rm{main}}$ is a stack on which $\tilde\pi_*\tilde f^*\mathcal O_{\PP^r}(l)$ is a vector bundle, which allows them to define a notion of \emph{reduced invariants} for complete intersections in $\PP^r$, based on the hyperplane property:  %The idea is that blowing up successively the locus of degree zero elliptic curve with one rational tail (\emph{elliptic tails}), then with two (\emph{elliptic bridges}), etc. and taking the fiber product
%\bcd
%\widetilde{\mathcal M}_{1,n}(\PP^r,d)\ar[r]\ar[d] & \M{1}{n}{\PP^r}{d}\ar[d] \\
%\widetilde {\mathfrak M}_{1,n}^{\operatorname{wt}=d} \ar[r] & \mathfrak M_{1,n}^{\operatorname{wt}=d}
%\ecd
%the strict transform of the main component becomes smooth and meets the rest of the components in normal crossing fashion (the \'{e}tale local model being the union of two linear subspaces of different dimensions in affine space). Furthermore, the sheaves $\tilde\pi_*\tilde f^*\mathcal O_{\PP^r}(l)$ contain a vector subbundle of rank $dl$ supported on the strict transform of the main component. This paved the way to the definition of reduced invariants.
\begin{dfn}
Let $X$ be a complete intersection of degree $(l_1,\ldots,l_k)$ in $\PP^r$. The genus one reduced Gromov-Witten invariants of $X$ are defined by integrating against: $$[\widetilde{\pazocal M}_{1,n}(\PP^r,d)^{\rm{main}}]\cap c_{\rm{top}}(\tilde\pi_*\tilde f^*(\oplus_{i=1}^k\mathcal O_{\PP^r}(l_i))).$$
\end{dfn}

Recently there has been a different attempt to discard some of the unwanted components. Viscardi \cite{VISC} has introduced a series of \emph{alternate compactifications} of the moduli space of maps from elliptic curves, based on Smyth's work on the birational geometry of $\oM_{1,n}$. In Viscardi's space $\oM_{1,n}^{(m)}(X,\beta)$ maps that contract an elliptic curve with at most $m$ rational tails are destabilised, and replaced by maps from $l$-elliptic fold singularities with $l\leq m$.

The main result of this paper is that reduced invariants of the quintic three-fold can be recovered from Viscardi's alternate compactification.

\begin{thm}
Let $X$ be a generic smooth quintic three-fold. Then:
\[
N_1^{\rm{red}}(X,d)=N_1^{\rm{cusp}}(X,d)
\]
where:
\[N_1^{\rm{red}}(X,d):=\deg \left([\widetilde{\pazocal M}_{1}(\PP^4,d)^{\rm{main}}]\cap c_{\rm{top}}(\tilde\pi_*\tilde f^*(\mathcal O_{\PP^4}(5)))\right)\]
and
\[N_1^{\rm{cusp}}(X,d):=\deg [\Mone{1}{X}{d}]^{\rm{vir}}.
\]
\end{thm}

The insight for this fact comes from the following formula, the proof of which was first given by J. Li and A. Zinger \cite{LZ} in the symplectic category:

$$N_1(X,d)=N_1^{\rm{red}}(X,d)+\frac{1}{12}N_0(X,d).$$
Noticing that the genus $0$ contribution to this formula only comes from the boundary component with a single rational tail of degree $d$ suggests that removing such component would provide a more direct approach to reduced invariants. This is precisely what Viscardi's space of 1-stable maps does.

\textcolor{red}{Li-Zinger's formula was later proved in the algebro-geometric setting by H.-W. Chang and J. Li in \cite{CL}. This was only the final of a series of papers where they developed the theory of stable maps with $p$-fields and applied it together with the techniques of cosection-localised virtual classes \cite{KLcosection}. $p$-fields provide a way around the fact that we cannot control the geometry of $\oM_1(X,d)$ and there is no compatible perfect obstruction theory relative to the inclusion $\oM_1(X,d)\hookrightarrow \oM_1(\PP^4,d)$; indeed Chang and Li managed to include the unwanted part of the obstruction theory in the moduli space, and by these means endow a well-understood space (very similar to $\oM_1(\PP^4,d)$) with an obstruction theory counting genus $1$ curves in the quintic.}
\textcolor{blue}{To prove Li-Zinger formula in the algebro-geometric setting, one would like to understand how the virtual class of  $\oM_1(X,d)$ splits on the various component of the moduli space, but this seems to be unaccesible. To overcome this difficulty, Chang Li developed the theory of stable maps with $p$- fields.}

The proof of our main theorem follows from adapting the techniques of Chang and Li to Viscardi's moduli spaces. We show the following:
\begin{prop}
There exists a well-defined $1$-stabilisation morphism at the level of weighted pre-stable curves:
\[
\MM_{1,n}^{\rm{wt}}\to \MM_{1,n}^{(1),\rm{wt}}
\]
\end{prop}
The fiber product
\[
\MM_{1,n}^{\rm{wt}}\times_{\MM_{1,n}^{(1),\rm{wt}}}\Mone{1}{\PP^4}{d}^p
\]
is endowed with a virtual class counting $1$-stable maps to the quintic three-fold (the arguments in \cite{CLpfields} carry over almost unchanged). On the other hand we find a way to compare the main contribution of this space with Vakil-Zinger reduced invariants. A dimensional computation shows that this is the only enumeratively meaningful component.

\subsection*{Outline of the paper}
maybe table of notations
\section{Genus one stable maps to $\PP^r$ - Equations, components and alternate compactifications}

\subsection{Local equations and components}
We start by recalling a description of the global geometry of $\M{1}{n}{\PP^r}{d}$. Besides the main component, which we have defined in the Introduction, for every positive integer $k$ and partitions $\lambda\vdash d$ into $k$ parts and $\mu\vdash n$ into $k$ parts such that no $(\lambda_i,\mu_i)_{i=1}^k$ is either $(0,1)$ or $(0,2)$, there is a \emph{boundary} irreducible component $D_{\lambda,\mu}(\PP^r,d)$ defined to be the closure of the locus where:
\begin{description}
\item[i] the source curve is obtained by gluing a smooth elliptic curve with $k$ many $\PP^1$'s as rational tails,
\item[ii] the map contracts the core elliptic curve to a point, and
\item[iii] the rational tail $R_i$ has degree $\lambda_i$ and $\mu_i$-many marked points.
\end{description}
Indeed $D_{\lambda,\mu}(\PP^r,d)$ is the image of the gluing morphism from the fiber product:
$$\M{1}{k}{\PP^r}{0}\times_{\PP^r}\prod_{i=1}^k\M{0}{\mu_i+1}{\PP^r}{\lambda_i}.$$

The minimal arithmetic genus $1$ subcurve is called the \emph{core}. We then have:
\begin{prop}\label{prop:components}
\emph{(1)} The ones that we have described above are all the irreducible components of $\M{1}{n}{\PP^r}{d}$, %The set of irreducible components of $\M{1}{n}{\PP^r}{d}$ is in bijection with $\{\text{stable}\ (\lambda\vdash d,\mu\vdash n)|\ \#(\text{parts})(\lambda)=\#(\text{parts})(\mu)\}\cup\{\rm{main}\}$,
i.e.
$$\M{1}{n}{\PP^r}{d}=\M{1}{n}{\PP^r}{d}^{\rm{main}}\cup\bigcup_{\lambda,\mu}D_{\lambda,\mu}(\PP^r,d).$$

\emph{(2)} A map $[f]$ lies in the boundary of the main component  if and only if:
\begin{itemize}
\item either $f$ is non-constant on an irreducible component of the core,
\item or if $(E\ {}_{\mathbf p}\!\sqcup_{\mathbf q}\bigsqcup_{i=1}^k R_i,f)$ is a stable map with $k$ rational tails, where $E$ is the maximal contracted genus one subcurve, then $\{\operatorname{d}\!f(T_{q_i}R_i)\}_{i=1}^k$ are \emph{linearly dependent} in $T_{f(E)}\PP^r$.
\end{itemize}
\end{prop}

This is due to R. Vakil and A. Zinger in \cite{Vre,VZpreview}. We shall later discuss a different proof of this fact based on local equations for the moduli space.

%In order to understand how the Gromov-Witten invariants differ from the naive expectation of counting smooth embedded genus $1$ curves in the given homology class, one needs a good understanding of the virtual fundamental class and of its splitting on the various components. 
%This has led to the study of explicit local equation for the moduli space of stable maps in a smooth (relative to a moduli space of decorated curves) ambient space \cite{HL}, which we shall now briefly explain.
We now review Hu-Li's procedure \cite{HL} to find local equations of $\M{1}{n}{\PP^r}{d}$ in a smooth ambient space. These are useful in describing the structure of the intrinsic normal cone and its splitting.

Recall that a map $C\to\PP^r$ is given by a line bundle $L$ on $C$ together with $r+1$ sections in $H^0(C,L)$ that generate the line bundle. It is therefore natural to embed the space of stable maps as an open inside $\pi_*\mathcal L^{\oplus r+1}$ on the universal Picard stack $\mathfrak{P}_{1,n}=\mathfrak{Pic}(\cC_{1,n}\to\MM_{1,n})$.
Hu and Li actually work on a local chart of the algebraic stack $\mathfrak M_{1,n}^{\rm{div}}$, parametrising families of pre-stable curves with a relative Cartier divisor of degree $d$. 
 %The problem is that this cone is not smooth, due to the jumps in the dimension of the fiber at the boundary of the moduli space.
%(besides there is a small issue due to the $\mathbb G_m$ automorphism group of objects of $\mathfrak{P}_{1,n}$, which can be avoided by dishomogenising).

Let $[f\colon C\to \PP^r]$ be a point of $\M{1}{n}{\PP^r}{d}$: we may fix homogeneous coordinates on $\PP^r$ in such a way that $D_0=f^{-1}\{x_0=0\}$ is a simple smooth divisor on $C$ (i.e. a $d$-uple of smooth points); this property will then hold in a neighbourhood of $[f]$. This gives a map from an \'{e}tale chart $\pazocal U$ of $\M{1}{n}{\PP^r}{d}$ to an \'{e}tale chart $\pazocal V$ of $\mathfrak M_{1,n}^{\rm{div}}$, that we may assume to land in the locus where the divisor is made of $d$ distinct smooth points. Notice that this locus is smooth over $\mathfrak M_{1,n}$ by the deformation theory of smooth subschemes. A morphism to $\PP^r$ shall now be thought of as a curve-divisor pair $(C,D)$ together with $r$ sections of $\mathcal O_C(D)$: the morphism can be written as $[1:u_1:\ldots:u_r]$, where $1$ means the given section of $\mathcal O_C\to\mathcal O_C(D)$.

%is the image of the constant global section along the given map $\mathcal O_C\to\mathcal O_C(D)$ (i.e. we have taken dishomogenised local coordinates on $\PP^r$).

Furthermore, \'{e}tale locally on $\M{1}{n}{\PP^r}{d}$, we may pick extra sections $\mathcal A$ and $\mathcal B$ of $\mathcal C_{1,n}(\PP^r,d)$ such that 
\begin{enumerate}
\item they pass through the core elliptic curve;
\item they are distinct smooth points disjoint from the support of $\pazocal D_0$. 
\end{enumerate}
On  $\pazocal V$ we have that  $\pi_*\OO_{\cC}(\pazocal D+\mathcal A)$ is a vector bundle and $\pi_*\OO_{\cC}(\pazocal D)$ is carved inside it by requiring that the restriction map $\pi_*\OO_{\cC}(\pazocal D+\mathcal A)\to\pi_*\OO_{\cC}(\pazocal D+\mathcal A)\rvert_{\mathcal A}$ is zero.

\begin{prop}\label{prop:smoothability}
\emph{(1)} \'{E}tale locally on $\M{1}{n}{\PP^r}{d}$, there is a locally closed embedding of $\M{1}{n}{\PP^r}{d}$ inside the vector bundle $V:=\Spec(\pi_*\OO_{\cC}(\pazocal D+\A)^{\oplus r})$ over $\pazocal V$:
\bcd
\M{1}{n}{\PP^r}{d} & \pazocal U\ar[l,"\acute{e}t"]\ar[r,hook]\ar[d] & V\ar[dl] \\
\MM_{1,n}^{\rm{div}} & \pazocal V\ar[l,"\acute{e}t"] &
\ecd
Notice that $V$ is smooth.
%It is obtained by imposing the nondegeneracy and stability conditions \emph{(open)} on the vanishing locus of the restriction of the universal section of $\pi_*\mathcal L(\A)^{\oplus r}$ to $\A$ \emph{(closed)}.

\emph{(2)} Assume furthermore that $\pazocal V$ is affine; then on $\pazocal V$ we have 
\[\pi_*\OO_{\cC}(\pazocal D+\A)\cong \pi_*\OO_{\cC}(\pazocal D+\A-\B)\oplus \pi_*\OO_{\cC}(\pazocal D+\A)_{|\B}\] %(trivially, $\mathcal O_{\mathcal V}^{\oplus d+1}\cong\mathcal O_{\mathcal V}^{\oplus d}\oplus\mathcal O_{\mathcal V}$)
and the restriction-to-$\A$ map is zero on the second factor.

\emph{(3)} Call $\varphi\colon \pi_*\OO_{\cC}(\pazocal D+\A-\B)\to\pi_*\OO_{\cC}(\pazocal D+\A)_{|\A}$ the map induced by restricting to $\A$.

 If we choose a suitable basis of $\pi_*\OO_{\cC}(\pazocal D+\A-\B)$, $\varphi$ can be written in a very explicit form: let $D=\sum_{i=1}^d\delta_i$, then  
 \[\varphi=\oplus\varphi_i\colon\oplus_{i=1}^d \pi_*\mathcal O_{\cC}(\delta_i+\A-\B)\to \pi_*\OO_{\cC}(\pazocal D+\A)_{|\A}.\] 
  Furthermore, after choosing trivializations for the line bundles above,  $\varphi_i\colon\mathcal O_\pazocal V\to\mathcal O_\pazocal V$ is given by multiplication by \[\prod_{q\in[\A,\delta_i]}\zeta_q,\]
   where $\zeta_q$ is the smoothing variable on $\mathfrak M_1$ corresponding to the node $q$, and $[\A,\delta_i]$ are all the nodes separating $\A$ (i.e. the core) from the point $\delta_i$.
\end{prop}
We refer the reader to \cite[Lemma~4.10, Proposition~4.13]{HL} for full details; we are going to review the key ideas in \S\ref{sec:equations}, where we adapt the theory to $\Mone{1}{\PP^r}{d}$. We include here a proof of Proposition~\ref{prop:smoothability}-(2) based on the local equations.

%Given these explicit equations and by carefully examining the proof of \cite[Prop. 4.13]{HL}, we can look back at the problem of determining the boundary of the main component of the space of stable maps. The following characterisation of smoothability was already known to (and partly proven in) \cite[Lem. 5.9]{Vre}.

%\begin{prop}\label{smoothability_with_eqn}
%Let $[f\colon C\to \PP^r]$ be a degenerate stable map from a ($n$-marked) genus one curve of the form $(E {}_{p_1,\ldots,p_m}\sqcup_{q_1,\ldots,q_m}\bigsqcup_{i=1}^m R_i,f)$, where $E$ is the maximal connected contracted arithmetic genus one subcurve, and $R_i$ are the rational tails (chains of $\PP^1$ along which $f$ has positive degree). The following are equivalent:
%\begin{enumerate}
%\item $[f]$ is smoothable;
%\item $\operatorname{d}\!f(T_{q_i}R_i)$ are linearly dependent in $T_{f(E)}\PP^r$;
%\end{enumerate}
%\end{prop}

\begin{proof}
 Let us start with the easiest degenerate situation: a contracted elliptic curve attached to a $\PP^1$ of degree $d$ at a single node $q$. Equations for the moduli space of maps around such a point look like 
 \[\zeta_q\sum_{i=1}^d w_i^j=0,\;\;\;\text{for} \;\;\;j=1,\ldots,r\]
 where the $w_i$ are coordinates on the fibers of $\pi_*\mathcal O_{\cC}(D+\A-\B)$ in the basis given by $\pi_*\mathcal O_{\cC}(\delta_i+\A-\B),\ i=1,\ldots,d$.
